---
title: "A worked example with `pcs`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Peter Sagan by the numbers

```{r, message = F}
library(dplyr)
library(stringr)
library(ggplot2)
library(pcs)
library(lubridate)

process <- F
# If you want to do a fresh download of Peter Sagan's record

if (process){
  # Extract race results and bio
  sagan <- query_pcs("Peter Sagan")
  sagan_results <- sagan$results
  sagan_bio <- sagan$profiles
  
} else {
  # Otherwise Sagan's results are saved as an example data set in the package (through 5/20/2021)
  sagan_results <- pcs::sagan$results
  sagan_bio <- pcs::sagan$profiles
}
```

#### How many professional races has Peter Sagan raced? How many has he won?

```{r}
results_finished <- sagan_results %>%  
  
  # Final placement in grand tour points classifications is included in 
  # the results data, so we filter them out first. We can also filter out
  # the races that Sagan entered but did not finish
  filter(!str_detect(stage, "classification"),
         result < 200) 

# Completed races
(races_finished <- nrow(results_finished))

# Appearances in the top 10
(top_10 <- results_finished %>% filter(result <= 10) %>% nrow())

# Professional wins
(wins <- results_finished %>% filter(result == 1) %>% nrow())
```

Since 2007, Peter Sagan has finished `r races_finished` races in total. Among those `r races_finished` races, he has placed in the top ten `r top_10` times, and has an incredible `r wins` wins to his name, giving him a historical winning percentage of `r round(wins/races_finished, 3)*100`%.

#### Visualizing Peter Sagan's career

There's little doubt that Peter Sagan is among the most winningest cyclists in history. Here's a visualization exploring Sagan's career arc; showing the frequency of races that he's finished in the top five by year.

```{r, message=F, warning=F}
results_finished %>% 
  dplyr::filter(result <= 5) %>% 
  group_by(year = year(date), result) %>%
  summarise(n = n()) %>%
  mutate(result = factor(result, levels = c(1:5))) %>% 
    ggplot() +
      geom_bar(aes(x = year, 
                   fill = result, y = n), 
               stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(y = "Number of finishes",
       fill = "Result") +
  theme(axis.title.x = element_blank())
```

