---
title: "PCS Usage"
author: "Sean Hardison"
date: "Latest version: `r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
params:
  EVAL: !r TRUE
vignette: >
  %\VignetteIndexEntry{PCS Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

```

## Introduction

The objective of the `pcs` package is to facilitate the analysis of professional cycling biographical data and rider results from [ProCyclingStats.com](https://procyclingstats.com). This vignette explores some of the ways that the data may be used. All biographical data is "as is" and may not be accurate beyond what is listed on ProCyclingStats. Here we show some worked examples making use of the `pcs` data sets. 

## Rider biographical data

### What countries are well represented in the peloton?

```{r, message=F, warning=F}
library(pcs)
library(ggplot2)
library(dplyr)

# Combine men and women rider profile data and count the number of riders
# by country
all_riders <- rider_profiles_men %>% 
  bind_rows(., rider_profiles_women) %>% 
  group_by(nationality) %>% 
  dplyr::summarise(n_riders = n()) %>% 
  top_n(n_riders, n= 25)

# Visualize
all_riders %>% 
  arrange(desc(n_riders)) %>% 
ggplot() +
  geom_bar(aes(x = reorder(nationality, -n_riders, sum),
               y = n_riders), stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(y = "Number of cyclists",
       x = "Country of origin") 
```

### Which countries produce the most successful climbing specialists?

```{r, message=F, warning=F}
climbers <- 
  rider_profiles_men %>% 
  mutate(sex = "men") %>% 
  bind_rows(., rider_profiles_women %>% 
              mutate(sex = "women"))

# Extract tenth percentile of climber points by sex
top_points_perc <- 
  climbers %>% 
  group_by(sex) %>% 
  summarise(q = quantile(climber, 
                         probs = seq(0,1,0.1)),
            prob = names(quantile(climber, 
                                  probs = seq(0,1,0.1)))) %>% 
  filter(prob == "90%") 

# Identify the countries that the most "prolific" climbers originate from. 
# Note that this does not correct for age of the rider. 
best_climbers <- 
  climbers %>% 
  left_join(.,top_points_perc, by = c("sex")) %>% 
  filter(climber >= q) %>% 
  group_by(sex, nationality) %>% 
  dplyr::summarise(n_riders = n()) %>% 
  top_n(n_riders, n = 10)

# Visualize
best_climbers %>% 
  arrange(desc(n_riders)) %>% 
ggplot() +
  geom_bar(aes(x = reorder(nationality, -n_riders, sum),
               y = n_riders), stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(y = "Number of cyclists",
       x = "Country of origin") 
```



